Toward Adaptive GPU Resource Management for Embedded Real-Time Systems

### 2012-03-10 draft/draft.tex

\section{Introduction}
\label{sec:introduction}

>>>>

\section{Introduction}
\label{sec:introduction}
Graphics processing units (GPUs) are widely used in a range of applications from supercomputers to mobile computers. For example, three of the top five supercomputers on the TOP500 list~\cite{TOP500} utilize GPUs to accelerate computations, and recent tablets such as ASUS Eee Pad Transformer Prime leverage 12 GPU cores of NVIDIA’s well-known SoC, Tegra 3 to boost the performance while saving power~\cite{Tegra3}. This trend will only continue and in fact is expected to accelerate across multiple application domains.

Modern automobiles have several tens of processing units, and advances in safe-driving features such as adaptive cruise control, stop-and-go cruise control, lane keeping and assisted lane change require even higher performance computing. As vehicles become semi-autonomous, drive autonomously on demand and eventually become fully autonomous, a multitude of computer vision, sensor fusion, signal processing and graphics problems must be solved in real time. Since solutions to those problems can leverage benefits of parallel processing techniques, GPUs can be used to accelerate computations. As an initial step, NVIDIA GPUs will be used for infotainment systems in future BMW vehicles~\cite{NVIDIA_BMW}.

Recent trends in self-driving cars~\cite{Urmson08, Markoff10, Kelly12} are motivated by various benefits of automated driving, but most importantly enabling automatic safety features. Those features are realized by the smart vehicle planning and the intelligent processing of data from various types of sensors on self-driving cars such as LIDAR (LIght Detection And Ranging), radar, camera, ultrasonic sensor, and so forth. One common characteristic of both the smart vehicle planning and the intelligent data processing is that their performances can be greatly accelerated using GPU, hence parallelism. For example, the more a self-driving car plans its potential paths to follow {\it a priori}, the better the quality of self-driving is. Since the self-driving car should follow only the best path, calculation of each potential path can happen parallel. A CMU team has already started using GPUs for autonomous driving computations. They have increased the performance of motion planning for autonomous driving up to 40 times using an NVIDIA GPU GTX 260 having 192 logical cores~\cite{McNaughton11}.

The perception system which intelligently processes the data from various sensors can also obtain benefits using GPU. The perception system of a self-driving car should be able to detect, classify and track the obstacles around itself. Various types of sensors will generate huge amount of information to understand the surroundings. In order to process the data, a parallel processing technique is of vital importance. For example, a new self-driving car at CMU manages 1536 objects from LIDAR sensors before they are fused with other types of sensor data. There has been on-going research using GPU to build a perception system~\cite{Ferreira11}, and their GPU implementation yielded 30,000-times-faster performance compared to the case of using only CPU.

As described above, there has been research on applying GPU to different applications on self-driving cars, and it is clear that GPU can provide great benefits on realizing self-driving car technologies. However, not much research has been done when those technologies are deployed together on self-driving cars, where the loads of each application dynamically vary depending on environments. The period and the computation time of the planning algorithms for autonomous driving highly depend on the vehicle speed, so the planning algorithms can be heavily loaded when the car is on highway. The load of the perception algorithms mainly depend on the number of obstacles around the car hence the perception system requires more computing resources when the car is driving in the urban area. Therefore, an intelligent method of sharing many cores on GPU would be essential when we use GPU on self-driving cars. For example, if a self-driving car has a 96-core GPU, the planning algorithm of the car can use 72 cores on the highway and use 12 cores in the urban environment. A self-driving car~\cite{Urmson08} require tens of tasks, and the dynamic core management should be fulfilled across all tasks if those tasks utilize GPU.

In this paper, we propose conceptual frameworks for GPU applications to adjust their task execution times depending on their given workloads.

The rest of paper is organized as follows. Section~\ref{sec:system_model} describes how our proposed system is modeled. Section~\ref{sec:adaptivity_support} presents the methods for adaptively managing the given GPU resources, and we conclude our paper in Section~\ref{sec:summary}.
